DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-04-17 paper/evaluation-cachepart.tex
Section 4.F revision

\begin{figure}
  \centering
  \includegraphics[width=.5\textwidth]{figs/cache-mapping}
  \caption{A memory mapping of the cache sets for the Broadcom 
BCM2837 processor used in the Raspberry Pi 3 Model B.}
  \label{fig:cache-mapping}
\end{figure}

@@

In order to validate the effectiveness of cache partitioning, we patch
the kernel with PALLOC~\cite{yun2014rtas}, which is a kernel-level
color-aware page allocator, supporting cache partitioning. 
Figure~\ref{fig:cache-mapping} shows the physical address mapping of
Raspberry Pi3's BCM2837 processor. We use the bits 13 and 14 for
coloring, which results in 4 page colors.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/palloc_multicore}
  \caption{Cache space sensitivity of the control loop exectuion time.}
  \label{fig:palloc_multicore}
\end{figure}

In the first experiment, we investigate the cache space sensitivity of
the DeepPicar's DNN-based control loop. Using PALLOC, we create 4
different cgroups which are configured to use 1, 2,
3, and 4 colors (25\%, 50\%, 75\%, and 100\% of the L2 cache
space, respectively). We then execute the control loop on each cache
partition (cgroup) and measure the average processing time.

If performance improves as a result of partitioning the shared cache 
then we know that the DNN, to some extent, relies on the shared cache 
in addition to the shared DRAM. On the other 
hand, if there is no improvement in performance, then it can be 
observed that the shared memory is more critical for the DNN.
%% We run the DNN model with different amounts of L2 cache being
%% available based on the number of colors being used. For example, if two 
%% colors are used, then 25\% of the L2 cache is available, whereas if all 
%% 8 colors are used, 100\% of the L2 cache is available. Particularly, we 
%% test the performance of the model when it utilizes a single core and
%% all four of the available cores. 
The results can be seen in 
\ref{fig:palloc_multicore}. As expected, the amount of L2 cache 
available to the model does not result in any significant changes in 
performance. 
%% This can also be seen by the L2 miss rates of the model 
%% seen in \ref{fig:palloc_multicore_l2missrate}. As more cache space 
%% becomes avaiable, the percentage of L2 misses decreases, which doesn't 
%% affect the real-time performance of the DNN. As such, we find that a 
%% single model is not sensitive to the shared L2 cache.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/palloc_bandwidth_exectime}
  \caption{Timing impact of co-scheduling memory intensive co-runners 
when cache partitioning is enabled.}
  \label{fig:palloc_bandwidth_exectime}
\end{figure}

We also run a single model alongside three memory intensive read co-runners 
again, but ensure that each task is given equal amounts of L2 cache space.
For this, we assign one color to each task. The results can be seen in Figure
\ref{fig:palloc_bandwidth_exectime}. When cache partitioning is enabled, 
DNN performance becomes slightly better, with a improvement of 
\textasciitilde6.5\%. However, as the performance improvement is minimal, 
we do not believe that cache partitioning is effective for protecting the model 
from interference from other memory intensive processes.
%% At the same time, though, the L2 miss rates of the model don't correlate 
%% to the timing increases seen, as is shown in 
%% \ref{fig:palloc_bandwidth_l2missrate}. Instead, we find that the number
%% of bus accesses increases when cache partitions are used, and they mirror
%% the timing increases of the model. This can be seen in Figure 
%% \ref{fig:palloc_bandwidth_bus}. These additional bus accesses are
%% most likely caused by the L2 prefetcher, as it can generate additional bus
%% accesses that aren't counted as L2 misses. As a result, we find that cache
%% partitioning is detrimental to the performance of the DNN so long as the 
%% 0L2 prefetcher is enabled (we have not yet found a way to disable it on
%% the Pi 3). Furthermore, no cache sensitivity was displayed by the model,
%% thus showing that the performance of the L2 cache is not vital for DNN 
%% performance when co-runners are introduced.

By partitioning the shared L2 cache, we find that no noticable 
improvements are gained and that performance remains consistent. 
In all experiements, the model shows no sensitivity to the shared L2 
cache. As a result, we conclude that cache partitioning is not an 
effective isolation mechanism, and that the performance of the shared 
DRAM controller is of greater importance to the real-time efficiacy of 
the DeepPicar.

>>>>

@@

Cache partitioning is a well-known technique to improve isolation in
real-time systems by giving a dedicated cache space to individaul task
or core. In this experiment, we use PALLOC~\cite{yun2014rtas}, a
page-coloring based cache partitioning solution for Linux, to
investigate the effect of cache partitioning in protecting DeepPicar's
DNN based controller.

\begin{figure}
  \centering
  \includegraphics[width=.5\textwidth]{figs/cache-mapping}
  \caption{Physical address mapping of L1/L2 caches of Broadcom
    BCM2837 processor in Raspberry Pi 3.}
  \label{fig:cache-mapping}
\end{figure}

Page coloring is a OS technique that controls the physical addresses
of memory pages. By allocating pages over non-overlapping sets of the
cache, the OS can effectively partition the cache.
Figure~\ref{fig:cache-mapping} shows the physical address
mapping of Raspberry Pi 3's BCM2837 processor, which has 32K private
L1 I\&D (4way) caches and a share 512KB L2 (16 way) cache. In order not
to partition the private L1 caches, we use the bit 13 and 14 for
coloring, which results in 4 usuable page colors.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/palloc_multicore}
  \caption{Cache space sensitivity of the DNN controller. \fixme{I
      think 'no part.' bar is not necessary here.}}
  \label{fig:palloc_multicore}
\end{figure}

In the first experiment, we investigate the cache space sensitivity of
DeepPicar's DNN-based control loop. Using PALLOC, we create 4
different cgroups which are configured to use 4, 3,
2, and 1 colors (100\%, 75\%, 50\% and 25\% of the L2 cache
space, respectively). We then execute the DNN control loop (inference)
on each cache partition (cgroup) and measure the average processing
time.

Figure~\ref{fig:palloc_multicore} shows the result. As can be seen,
the DNN inference timing is hardly changed at all regardless of the
size of the allocated L2 cache space. In other words, we find that
the DNN workload is largely insensitive to L2 cache space.

The next experiment further validates this finding. In this
experiment, we repeat the experiment in
Section~\ref{sec:eval-memhog}---i.e., co-scheduling the DNN model and
three Bandwidth (BwRead or BwWrite) instances---but this time we
ensure that each task is given equal amounts of L2 cache space by
assigning one color to each task's cache partition.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/palloc_bandwidth_exectime}
  \caption{Average processing time vs. the number of memory
memory intensive co-runners. Unlike
Figure~\ref{fig:perf_vs_bandwidth}, the shared L2 cache is equally
partitioned and each task is given a dedicated cache
partition. \fixme{how about changing the figure style to be the same
  as Figure 9.}}
  \label{fig:palloc_bandwidth_exectime}
\end{figure}

Figure \ref{fig:palloc_bandwidth_exectime} shows the
result. For BwRead, when cache partitioning is enabled, DNN
performance becomes slightly better, with a improvement of  
\textasciitilde6.5\%. For BwWrite, cache partitioning actually makes
the DNN performance slightly worse that that of w/o partitioning. In
either case, the main observation is that cache partitioning is not
effective in providing timing isolation to our DNN workload.

%% By partitioning the shared L2 cache, we find that no noticable 
%% improvements are gained and that performance remains consistent. 
%% In all experiements, the model shows no sensitivity to the shared L2 
%% cache. As a result, we conclude that cache partitioning is not an 
%% effective isolation mechanism, and that the performance of the shared 
%% DRAM controller is of greater importance to the real-time efficiacy of 
%% the DeepPicar.

%% If performance improves as a result of partitioning the shared cache 
%% then we know that the DNN, to some extent, relies on the shared cache 
%% in addition to the shared DRAM. On the other 
%% hand, if there is no improvement in performance, then it can be 
%% observed that the shared memory is more critical for the DNN.

%% We run the DNN model with different amounts of L2 cache being
%% available based on the number of colors being used. For example, if two 
%% colors are used, then 25\% of the L2 cache is available, whereas if all 
%% 8 colors are used, 100\% of the L2 cache is available. Particularly, we 
%% test the performance of the model when it utilizes a single core and
%% all four of the available cores.

%% As expected, the amount of L2 cache 
%% available to the model does not result in any significant changes in 
%% performance. 
%% This can also be seen by the L2 miss rates of the model 
%% seen in \ref{fig:palloc_multicore_l2missrate}. As more cache space 
%% becomes avaiable, the percentage of L2 misses decreases, which doesn't 
%% affect the real-time performance of the DNN. As such, we find that a 
%% single model is not sensitive to the shared L2 cache.

@@

%% At the same time, though, the L2 miss rates of the model don't correlate 
%% to the timing increases seen, as is shown in 
%% \ref{fig:palloc_bandwidth_l2missrate}. Instead, we find that the number
%% of bus accesses increases when cache partitions are used, and they mirror
%% the timing increases of the model. This can be seen in Figure 
%% \ref{fig:palloc_bandwidth_bus}. These additional bus accesses are
%% most likely caused by the L2 prefetcher, as it can generate additional bus
%% accesses that aren't counted as L2 misses. As a result, we find that cache
%% partitioning is detrimental to the performance of the DNN so long as the 
%% 0L2 prefetcher is enabled (we have not yet found a way to disable it on
%% the Pi 3). Furthermore, no cache sensitivity was displayed by the model,
%% thus showing that the performance of the L2 cache is not vital for DNN 
%% performance when co-runners are introduced.

### 2018-04-17 paper/evaluation.tex
Section 4.F revision

Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure, the
X-axis shows the system configuration: \#of DNN models x \#of CPU
cores/DNN. For example, '4Nx1C' means running four DNN models each of
which is assigned to run on one CPU core, whereas `2Nx2C' means running
two DNN models, each of which is assigned to run on two CPU
cores. The Y-axis shows the average inference timing.
The two bars on the left show the impact of co-scheduling four DNN
models. Compared to executing a single DNN model on one CPU core
(1Nx1C), when four DNN models are co-scheduled (4Nx1C), each model
suffers an average inference time increase of approximately 11 ms,
$\sim$24\%. On the other hand, when two DNN models, each using two CPU
cores, are co-scheduled (2Nx2C), the average inference timing is increased by
about 4 ms, or 13\%, compared to the baseline of running one model
using two CPU cores (1Nx2C).

>>>>

Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure, the
X-axis shows the system configuration: \#of DNN models x \#of CPU
cores/DNN. For example, `4Nx1C' means running four DNN models each of
which is assigned to run on one CPU core, whereas `2Nx2C' means running
two DNN models, each of which is assigned to run on two CPU
cores. The Y-axis shows the average inference timing.
The two bars on the left show the impact of co-scheduling four DNN
models. Compared to executing a single DNN model on one CPU core
(1Nx1C), when four DNN models are co-scheduled (4Nx1C), each model
suffers an average inference time increase of approximately 11 ms,
or 24\%. On the other hand, when two DNN models, each using two CPU
cores, are co-scheduled (2Nx2C), the average inference timing is increased by
about 4 ms, or 13\%, compared to the baseline of running one model
using two CPU cores (1Nx2C).
